### 2023-09-20 corridor_docs/thesis_ideas.Rmd
adding thesis ideas doc

---
title: "Project Ideas and Notes"
author: "Eric Escobar-Chena and Emilio M. Bruna"
date: "updated: `r Sys.Date()`"
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
# - \fancyfoot[R]{last update `r Sys.Date()`}
- \fancyfoot[C]{ }
- \fancyhead[R]{p. \thepage}
- \fancyhead[L]{last update `r Sys.Date()`}
- \usepackage[default]{sourcesanspro}
# - \usepackage{parskip}
# - \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
# - \usepackage{sectsty} \sectionfont{\centering}
# - \usepackage{titlesec}
# - \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
# - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
bibliography: "eric-thesis.bib"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(citr)
```

### _Notes from conversation with Julian:_

1. Low-hanging fruit: Effect of connectivity on (a) species and functional group diversity and abundance
1. Next level: Movement & Dispersal. There are large areas of forest where it is possible to capture beetles for Mark-Release-Recapture (MRR) experiments.
1. Higher-risk, Higher reward (even if they don't go in MS, can be set-up for potential PHD projects): Experiments on Ecosystem Services

- dung burial, decomposition, soil properties based on results of 1 & 2 (buckets with dung and beetle assemblages)
- gas (e.g., methane) emissions
- seed dispersal/burial/germination

### Example studies

1. Diversity and Abundance

+ @WOS:000178929600002: "56% of individuals were capturedin the continuous forest, 29% in the mosaic habitat and 15% in the forest fragments"

2. Movement & Dispersal

3. Ecosystem Services

# Other stuff to work on:

1. Species List & Keys for Dung Beetles of the Southeastern US
1. Any previous work done on dung beetles in Southeastern US
1. List of Equipment and Tools needed
+ Dung beetle traps
+ alcohol
+ bait

References