### 2025-02-25 corridor_docs/eric_ms_thesis/thesis_planning.Rmd
added methoods

---
title: "Influence of Connectivity on Dung Beetle Communities"
author: "ERIC IN ALL CAPS"

# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
 - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
# # - \fancyfoot[R]{ }
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  pdf_document:
    includes:
      # in_header: style/preamble.tex
      template: null
      keep_tex: no
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

# Abstract

Thesis abstracts should be 250 words or less.

\newpage

# INTRODUCTION

INTRODUCTION

Corridors have been shown to be an important mechanism for facilitating the movement of organisms through fragmented landscapes [@haddad_corridor_2003]. These movements are hypothesized to prevent species diversity from declining in fragments, as well as help maintain the ecosystem services provided by these species (at both the patch- and landscape-level)[@burt_ants_2022]. Although there is some evidence that animals disperse between patches via corridors, and that connected patches have higher species diversity than unconnected ones, little work to date has investigated the consequences of these corridor-driven patterns for ecosystem services.

 Dung beetles have emerged as a model system with which to test hypotheses on how changes in landscape structure driven by human activities influence biodiversity and their ecosystem services [@ros_how_2012; @roslin_dung_2000]. The removal, breakdown, and burial of animal feces is an important ecosystem service provided by dung beetles such as enhanced nutrient cycling and soil quality and reduction of parasites on methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]. Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits (e.g., size, foraging style, resource-use) [@WOS:000891747700001]. Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001]. However, it remains unknown if corridors actually act to reduce the loss of dung beetle species from fragments, if such declines are influenced by inter-specific differences in dispersal capability, and what the consequences of these patterns are for the ecosystems services they provide. One major factor behind this lack of information is the challenge in finding locations where one can assess the role of corridors while also while controlling for confounding factors such as patch size, edge, and corridor length [@haddad_habitat_2015].

 **We sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches.

- First Paragraph: What is the topic of your introduction and why is it important/interesting/relevant?

What?

Connectivity and fragmentation

understanding how movement corridors effect populations dung beetles

Dispersal ability

Why?

Understanding responses to landscape change allows us to anticipate how environments might change with further disturbance.

Patterns in population distributions might suggest how we can study focal organisms in the future.

Could help understand how interactions may be changing.

- Second Paragraph: What is known about this topic already?

Effects of landscape change are already very well studied, accross different taxa

effects of edge, patch size, and direct habitat loss

using corridors to connect fragmented landscapes is also well studied and for many taxa

dung beetles as a model system for studying movement, functional diversity, and ecosystem health.

dung beetle populations in connected and fragmented habitats

insect declines in fragmented landscapes

studying coprophagous insects

- Third Paragraph: What isn't known about this topic and why might it change how we think/act about the topic?

How corridors directly impact dung beetle communitites

are changes driven by dispersal ability

at our study site we can directly compare patch connectivity and patch shape with fragmented landscapes to obtain a strong idea of what landscape features effect dung beetle collection.

- Fourth Paragraph: Why hasn't this thing been studied/assessed/done before?

other work at the corridor project but we are doing dung beetles instead

- Fifth Paragraph: Literally the words "Here we..."

Here sampled the commmunity of of dung beetles at the SRS Corridor Experiment to test the following prediction:

Species richness, diversity, and functional diversity will be higher in patches connected by a movement corridor than in patches that are unconnected.

# Methods

Study site

- description of srs

Our study took place at the Savannah River Site(SRS), a National Environmental Research Park in southern South Carolina, US(33.208 N, 81.408 W) in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals(Tewksbury et al. 2002). Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna. In each replicant the central patch (100 x 100 m) is always connected to one peripheral patch with identical dimensions by a 150 x 25 m corridor, this will hereafter be referred  to as the connected patch. The remaining patches are either "winged" or "rectangular".  The winged patch is also 100 x 100 m, however they exhibit their characteristic wings in the form of two 75 x 25 m offshoots meant to account for the extra area and edge space the corridor provides. The rectangular patch is 100 x 137.5 m also the same area as the space of the connected patch plus the corridor. Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch. For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.

Dung beetle sampling

In the months of July and August 2024 dung beetles were sampled in 4 blocks spread across SRS, baited pitfall traps were placed in one of each patch type and in one matrix plot per block. Traps were placed in groups of 3 in the centers of each patch approximately 250 meters from the midpoint of the central patch 40 m from patch edge. Pitfalls were oriented in a triangular pattern with the bottom two traps positioned towards the center patch, each trap 20 m apart. Plots in the matrix were set up in a similar fashion with the center point 250 m from the center placed equidistant between adjacent patches.  For each sample period, traps were baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing. In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected. All dung beetles were counted and identified to species with the exception of beetles of the genus Aphodius, overall 15 species were identified and approximately 5300 individual beetles were collected.

## Study site

  - description of srs
  - experimental design 
  - conditions during sample period
  - historical significance of site and experimental design 
  - justification for selected patches

## Dung beetle sampling

  - structure and arrangment of traps 
  - description of traps
  - bait 
  - sample period
  - ID 
  - biomass if we do biomass

### Analyses

1. **Species Richness:** absolute number? non-parametric estimators?
  - Michaelis-Menten
  - choa?
1. **Species Diversity:** what index should we compare?
  - alpha diversity per patch type
  - beta between patch types
  - hill numbers

1. **Functional Diversity:** Need to assign each species to a functional group: roller, tunneler. dweller, others? 
  - habitat preference (forest, pasture, generalist)

>_Look through dung beetle pubs and see how/what people compare_

  - lets hammer this out 
  - modeling? 
    - glmm with poisson dist reccomended by julian
  - beta, abundance, biomass? per site
  - species list by sampling blocks (anything with this?)
  - habitat preference 
  - rarefaction

# RESULTS

some summary statistics:

1. total number of beetles _from_ total number of species 
1. were all species found in all habitats? Were any species found in only 1 habitat?
1. Were all species found in all blocks? Were any restricted to only 1-2 blocks?
1. Were all species found in the matrix? (expect that so, since it is the 'baseline' or 'source' habitat)
1. Number of species in each functional group
1. most common 3-4 species
1. any rare species?

# DISCUSSION

1. dont forget o discuss the basic biology...why might a species be so common? why might one be rare?

# ACKNOWLEDGMENTS

Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents. The heading “ACKNOWLEDGMENTS” uses the 002 CHAPTER TITLE style.. The paragraphs in this section should use the style called 006 Body Text

# OTHER REQUIRED TEXT

### Dedication 
Your dedication is typed here. It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally

### List of Abbreviations
1. A word to be defined:	Write the definition here. 
1. Another word:	And the list continues with another definition.

# Biographical Sketch
A biographical sketch is required of all candidates. The biographical sketch should be in narrative form. Third person, past tense, it typically includes the educational background of the candidate. The author should have replaced this paragraph with their own.