### 2025-07-10 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
update and formatting edits

---
title: "The Influence of Connectivity on Dung Beetle Communities"
author: "ERIC ESCOBAR-CHENA"
# date: "\\vspace{-1em}"
# header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
#  - \usepackage{lineno}
#  - \linenumbers
#  - \usepackage[default]{sourcesanspro}
#  - \usepackage{fancyhdr} # modify header into left, center, right
#  - \pagestyle{fancy} # cues the use of fancy headers
#  - \fancyfoot{} # cues the use of fancy footer
#  - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
#  - \fancyhead[R]{p. \thepage} # name and page on header right
#  - \usepackage{subcaption}
#  - \usepackage{float}
# # # - \fancyfoot[R]{ }
#  - \usepackage{setspace}

header-includes:
  - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
  - \usepackage{float} #to keep figure in the location placed instead of end
  - \floatplacement{figure}{H}
  - \usepackage{booktabs}
  - \raggedbottom
  - \usepackage{fancyhdr}
  - \pagestyle{plain}
  - \fancyfoot[C]{\thepage}
  - \fancyhead[L]{}
  - \fancyhead[R]{\small{EEC MS Thesis Draft}}
  - \thispagestyle{fancy}
  - \usepackage{tabu}
  - \usepackage{makecell}
  - \usepackage{pdflscape}
  - \usepackage{longtable}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \setlength\parindent{22pt}
  - \usepackage{indentfirst}
  # - \usepackage{caption}
  # - \captionsetup[figure]{name={Figure 1-},labelsep=period}
  # - \captionsetup[table]{name={Table 1-},labelsep=period}
  - \renewenvironment{abstract}{}{} #suppress the abstract
  - \usepackage{setspace}
  - \setcounter{secnumdepth}{0} # removes numbering from sections and subsections
  - \usepackage{sectsty}\allsectionsfont{\raggedright}
  # - \usepackage{sectsty}\sectionfont{\centering}{\normalfont} # center # headings
  # - \usepackage{sectsty}\subsectionfont{\centering}{\normalfont} # center ## headings
  - \usepackage{sectsty}\subsubsectionfont{\bfseries} # center ## headings
  # - \usepackage{sectsty}\subsubsectionfont{\itshape} # no centering but italic subsu
  # - \usepackage{sectsty}\subsubsectionfont{\centering}{ital} # center ### headings
  # - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}

 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: true
    fig_caption: true
classoption: 12pt
# output: bookdown::word_document2
# output: word_document
# output: html_document

# output: bookdown::pdf_document2
  # pdf_document:
  #   includes:
  #     # in_header: style/preamble.tex
  #     template: null
  #     keep_tex: no
  #     extra_dependencies: ["float"]
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

```{r GlobalOptions, include = FALSE}
options(knitr.duplicate.label = 'allow')
options(knitr.graphics.error = FALSE)  # This allows for the include_graphics to not use the path from root 
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
knitr::opts_chunk$set(tab.pos = "H", out.extra = "")
```

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
library(here)
library(tidyverse)
library(kableExtra)
# library(grateful)
library(knitr)
# library(grid)
# library(gridExtra)
# library(png)
# library(egg)
# library(magick)
# library(citr)
# library(devtools)
# devtools::install_github("crsh/papaja")
# cite_packages(citation.style = "peerj", out.dir = ".")
```

```{r load-data, cached=TRUE, echo = FALSE, warning=FALSE, message = FALSE}

btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))

# total beetles captured 
total_N<-btl_data %>% 
  summarise(n=sum(sum, na.rm=TRUE))

# total no of spp
total_spp<-btl_data %>% 
  select(-sum) %>% 
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(desc(n)) %>%
  mutate(species = reorder(species, desc(n))) %>% 
  summarise(n_spp=n_distinct(species))

# 
# most_common<-spp_table_data %>% select(Species) %>% slice(1)
# most_common_n<-spp_table_data %>% select(N) %>% slice(1)

```
\newpage

\vspace*{5cm}

\begin{center}
THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
\end{center}

\vspace{5cm}

\begin{center}
By
\end{center}

\vspace{1cm}

\begin{center}
ERIC ESCOBAR-CHENA
\end{center}

\vspace{5cm}

\begin{center}
A THESIS PRESENTED TO THE GRADUATE SCHOOL\\  
OF THE UNIVERSITY OF FLORIDA IN PARTIAL FULFILLMENT\\ 
OF THE REQUIREMENTS FOR THE DEGREE OF\\  
MASTER OF SCIENCE\\
\end{center}

\begin{center}
UNIVERSITY OF FLORIDA
\end{center}

\begin{center}
2025
\end{center}

\newpage

\vspace*{11cm}

\begin{center}
© 2025 Eric Escobar-Chena
\end{center}

\newpage

\vspace*{10cm} 
<!-- _Your dedication is typed here. It should begin with the word “To.” (To my Mom is a typical dedication) If your dedication is longer than a single line it should be single-spaced, and centered vertically and horizontally_ -->

\begin{center}
To my mom, dad, and two sisters. This thesis is a chapter in our immigrant story. 
\end{center}

\newpage

\begin{center}

\textbf{ACKNOWLEDGMENTS}

\end{center}
<!-- _Acknowledgments must be written in complete sentences. Do not use direct address. For example, instead of Thanks, Mom and Dad!, you should say I thank my parents._  -->
\doublespacing

I thank the USDA Forest Service for maintaining experimental landscapes and assisting in getting established at the site. I also wanted to specifically thank Thomas Smith for his help in data collection, Sara Escobar-Chena for her help in processing and data entry.

\newpage

\begin{center}  
\textbf{LIST OF ABBREVIATIONS}
\end{center}

\vspace{1cm}

\begin{flushleft}

SRS: Savannah River Site.

GLMM: Generalized Linear Mixed Model.

\end{flushleft}

\newpage

\vspace*{5cm}

\singlespacing

\begin{center}  
Abstract of Thesis Presented to the Graduate School\\  
of the University of Florida in Partial Fulfillment of the\\  
Requirements for the Degree of Master of Science.\\   
\end{center}

\vspace*{1cm}

\begin{center}
THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES
\end{center}

\begin{center}
By
\end{center}

\begin{center}
Eric Escobar-Chena   
\end{center}

\begin{center}
August 2025   
\end{center}

\begin{flushleft}
Chair: Emilio Bruna\\  
Major: Wildlife Ecology and Conservation\\   
\end{flushleft}

\doublespacing

Habitat fragmentation threatens biodiversity across the globe as habitat loss, isolation, and edge effects become increasingly prevalent. Corridors have become an important tool in order to combat the negative effects of fragmentation, however they are difficult to study in natural systems without incurring confounding effects. To observe changes in insect community composition as an effect of landscape features I sampled dung beetles in a landscape scale experiment. I did not see a difference in species richness or diversity, but dung beetle abundances were higher in continuous forest habitat and open habitat patches connected by a corridors than in isolated patches.

\newpage

\begin{center}
CHAPTER 1\\  
THE INFLUENCE OF CONNECTIVITY ON DUNG BEETLE COMMUNITIES\\  
\end{center}

\doublespacing

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented [@diaz_pervasive_2019; @ma_global_2023; @taubert_global_2018]. Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003; @fletcher_is_2018]. However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012; @haddad_corridor_2003; @jennings_composition_2006; @harrison_habitat_1999]. This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions  [@haddad_habitat_2015].

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation [@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995]. The resulting changes in species composition are important to identify because any species impacted would have corresponding effects depending on how they interact with the ecosystem [@zhou_linking_2023]. Any gain or loss in key members of an community could disrupt processes which on their own could shape ecosystems [@WOS:000374962000009], or effect other organisms which rely on said interaction [@wu_brown-world_2011]. Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity [@debinski_survey_2000].

<!-- stuff that these inviduals or species do: -->

<!-- 1) their interactions with other species -->

<!-- - predation -->

<!-- - pollination -->

<!-- - seed dispersal -->

<!-- 2) stuff on their own that influences ecosystems -->

<!-- - ecosystem engineering (beavers, leaf-cutter ants) -->

<!-- - nutrient cycling -->

<!-- - decomposition -->

By measuring changes in biodiversity and species richness within experimental designs I am able to isolate factors might be contributing to ecological patterns and processes [@fletcher_jr_landscape_2023; @resasco_contribution_2017]. Past studies have endeavored to experimentally measure changes in community compositions as a result of connecting habitats with corridors [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002]  Yet very few have directly compared matrix and patch populations. Land use is different from one species to another so it is vital to understand where compositions are distributed and what processes might be driving population differences[@haddad_corridor_1999].

<!-- 1) We know far less on how corridors influence functional diversity. Functional diversity is important for the maintenance of ecosystems services, etc, so corridors cvould help ensure that.  (in context of dung beetles that would be feeding guilds (simplest), morphology etc. in deCastro Arrazola et al) -->

<!-- 2) Studies investigating corridors rarely look at whats i the matrix. The matrix could be a source of diversity that moves into patches and then through corridors or a completely different subset of species. (any other options?) Understanding whichh of these it is is important because it leads to differnt conclusions abiout role of corridors and conservation strategies. can also be spillover from matrix into patches and if depednent on corridors or edge -->

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions [@hasan_dung_2024]. The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013; @nichols_ecological_2008]. Local assemblages of dung beetles can be species-rich with species comprising a broad range of functional traits [e.g., size, foraging style, resource-use, [@decastro-arrazola_trait-based_2023; @ospina-garces_dung_2018]]. Previous studies have shown that isolated patches of habitat frequently have lower dung beetle diversity and abundance than areas of continuous habitat, as well as documented their presence in linear strips of habitat that resemble corridors [@WOS:000707546700001]. Past studies have also focused on how landscape structure alters the community compositions of dung beetles[@WOS:000398662200001], yet large landscape scale experimental studies with carefully controlled and replicated treatments are non-existent for this model species.

Here, I aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources [@hanski_dung_1991], interact with corridors in their landscapes. I sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch area to edge ratio and distance to edge [@tewksbury_corridors_2002]. I do so I address these  questions:

(1) How does the abundance of dung beetles differ among and between isolated and connected patches and how does this compare with abundance in matrix habitat?
(2) Does species richness and diversity differ between among connected patches, unconnected patches, and the matrix habitat?
(3) What are the implications of changes in community composition for ecosystem services and function?

<!--how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another? -->

<!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->

\begin{center} 
\textbf{Methods}
\end{center}
### Study site

My study took place at the Savannah River Site (SRS), a National Environmental Research Park in southern South Carolina, USA (33.208$^\circ$ N, 81.408$^\circ$ W, Figure \@ref(fig:srs-map)). in four of seven experimental landscapes designed for the purposes of directly observing the impacts of corridors and patch shape on the movements of plants and animals [@tewksbury_corridors_2002]. Each experimental landscape, termed blocks, consists of four patches of open habitat around a central patch all together within a matrix of pine savanna (Figure \@ref(fig:patches)). In each block the central patch (100 $\times$ 100 m) is always connected to one peripheral patch with identical dimensions by a 150 $\times$ 25 m corridor, this will hereafter be referred to as the connected patch. The remaining patches are either "winged" or "rectangular". The winged patch is also 100 $\times$ 100 m, however they exhibit their characteristic wings in the form of two 75 $\times$ 25 m offshoots meant to account for the extra area and edge space the corridor provides. The rectangular patch is 100 $\times$ 137.5 m also the same area as the space of the connected patch plus the corridor. Each block has a duplicate of either the winged or rectangle patch, all peripheral patches being 150 m from the center patch. For this study sampling was done in one of each patch type and in one matrix plot per block, all matrix blocks were set up 150 m away from the center as well.

### Dung beetle sampling

Dung beetles were sampled in July and August 2024 in four of the SRS blocks (8, 52, 53n, and 54; Figure \@ref(fig:srs-map)). In each block sampling was conducted using baited pitfall traps placed in each patch type as well as the matrix surrounding the patches (Figure \@ref(fig:pitfalls)). Traps were placed in groups of 3 in the centers of each patch, approximately 250 meters from the midpoint of the central patch 40 m from patch edge. Pitfalls were oriented in a triangular pattern with the bottom two traps positioned towards the center patch, each trap 20 m apart. Plots in the matrix were set up in a similar fashion with the center point 250 m from the center placed equidistant between adjacent patches. Individual pitfall traps consisted of two components, a 10cm tall by 8 cm wide cylinder base topped with a funnel with a 10cm wide rim. I sourced pig feces from the University of Florida Swine Barn Unit. Bait was processed into 5cm wide balls and wrapped in a layer of coffee filter material. For each sample period, traps were buried flush with the ground and baited with pig dung between 8-9 pm and picked up 12 hours later, all beetles captured were stored in ethanol for further processing. In total 16 sampling rounds were carried out with 4 rounds per block, 196 samples were collected.

All dung beetles were counted and identified to species using @nemes_illustrated_2015 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements. Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.

### Analyses

Biodiversity between patch types was compared using Hill numbers, a set of indexes developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006]. They are an alternative to more specialized metrics such as alpha, beta, and gamma diversities while being more standardized than other indexes such as Renyi or HCDT entropies, of which both groups of metrics are less intuitive for interpretation. Hill numbers are now the preferred metric for describing community dynamics for two reasons. First, they are extrapolated from the same equation, manipulating a single parameter (i.e., *q*) to arrive at estimates of richness and diversity. Second, by manipulating *q* we can gain an understanding of compositional shifts otherwise obscured while using species richness [@chao_rarefaction_2014]. I compared community composition by increasing magnitudes of diversity components (i.e., *qD*) of *0D* (i.e., species richness), *1D* (i.e., Shannon entropy), and *2D* (i.e, Simpson Diversity). Diversity numbers and species richness were calculated using the package `hill` [@li_hillr_2018] for the R statistical programming language [@posit_team_rstudio_2025]. Diversity numbers were calculated using package `iNEXT` [@hsieh_inext_2016]. Bray-Curtis dissimilarity values were calculated using package `Vegan` [@oksanen_vegan_2025]. Dung beetles were assigned traits by waste removal guild and habitat preference.

To test for the effects of connectivity on abundance, species richness, and species diversity I compared the values of the Hill Shannon and Simpson indexes in the different patch types and matrix. For abundance and richness I used generalized linear mixed models (i.e., GLMM) fitted to a poisson distribution [@bolker_generalized_2009]. Compared (1) the overall species richness and (2) the abundance of the top 6 most common species in each patch type. I included the identity of the sampling block as a random effects. To model my diversity metrics I took a similar approach, but this time using linear mixed effects models with a Gaussian distribution [@chao_rarefaction_2014]. In all models, the reference level for patch type was Matrix. All models were fit using `lme4` package [@bates_fitting_2015]. Prior to conducting my modeling I evaluated the the suitability of my data with *qqplots* generated with the `DARMa` package [@hartig_dharma_2024].

I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation. Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.

<!-- ### Functional diversity -->

<!-- 1. 1nd pass: roller, tunneler, dweller -->

<!-- 1. 2nd: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025) -->

<!-- 1. table: results of analyses  -->

<!-- 1. visualization: biomass of different (R/T/D) functional groups?    -->

<!-- _Reason for this breakdown: _ - richness and diversity are "community level" response. abundance is a species-level response. changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to functional group.  -->

<!-- Functional Diversity: Need to assign each species to a functional group: roller, tunneler. dweller, others?  - habitat preference (forest, pasture, generalist). Look through dung beetle pubs and see how/what people compare_ what stats / computer packages did you use to analyze and visualize the data? -->

\begin{center} 
\textbf{Results}
\end{center}

```{r data-load, include = FALSE, echo = TRUE, message = FALSE, warning = FALSE}

btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))
# total beetles captured --------------------------------------------------
total_N<-btl_data %>% 
  summarise(n=sum(sum, na.rm=TRUE))

## How many species total?
total_spp<-btl_data %>% 
  select(-sum) %>% 
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(desc(n)) %>%
  mutate(species = reorder(species, desc(n))) %>% 
  summarise(n_spp=n_distinct(species))

## Abund by patch

btl_sums_patch<-btl_data %>%
  select(-sum) %>%
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
  group_by(patch) %>%
  summarize(n=sum(n, na.rm=TRUE))

n_corr<-btl_sums_patch %>% filter(patch=="c") %>% select(n) %>% as.numeric()
n_mat<-btl_sums_patch %>% filter(patch=="m") %>% select(n) %>% as.numeric()
n_rect<-btl_sums_patch %>% filter(patch=="r") %>% select(n) %>% as.numeric()
n_wing<-btl_sums_patch %>% filter(patch=="w") %>% select(n) %>% as.numeric()

## Abundance by species
spp_abund<-btl_data %>%
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species) %>% 
  summarise(n=sum(n, na.rm=TRUE)) %>% 
  # filter(n>500) %>% 
  ungroup() %>% 
  select(species,n)%>% 
  arrange(desc(n)) %>% 
  rename("sp_code"="species") %>% 
  left_join(spp_codes) %>% 
  mutate(species=tolower(species)) %>% 
  mutate(species=paste(genus,species,sep=" ")) %>% 
  select(-c(subspecies,notes,ordway,srs,Guild,family,sp_code,genus))

## Common species

# by number (i.e., N most common)
how_many_common<-6
most_common_spp<-spp_abund %>% 
  slice_head(n = how_many_common)
count_most_common<-sum(most_common_spp$n)

# by abundance (i.e., over NNNN sampled)
high_abund_cutoff<-500
most_common_count<-spp_abund %>% 
  filter(n>high_abund_cutoff)

## Rare species

# by number (i.e., N most common)
how_many_rare<-4
rare_spp<-spp_abund %>% 
  slice_tail(n=how_many_rare)

# by abundance (i.e., over NNNN sampled)
low_abund_cutoff<-10
rare_spp_count<-spp_abund %>% 
  filter(n<low_abund_cutoff)

## Beetles by species and plot type (for Table 1)
btl_sums_patch<-btl_data %>%
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species,patch) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(patch,desc(n)) %>% 
  rename(sp_code=species,
         n_patch=n) %>% 
  pivot_wider(
    names_from = patch,
    values_from = c(n_patch)) %>% 
  left_join(spp_codes) %>% 
  mutate(species=tolower(species)) %>% 
  mutate(species=paste(genus,species,sep=" ")) %>% 
  ungroup() %>% 
  select(species,Guild,m,c,w,r) %>% 
  rename(Species=species)

# which ones NOT found in all habitats?

not_in_all_hab<-btl_sums_patch %>% 
  rename(species=Species) %>% 
  left_join(spp_abund) %>% 
  select(-n,-Guild) %>% 
  mutate(row_sums=m+c+r+w) %>% 
  filter(row_sums<4)

# row_sums==1 for any restricted to only one habitat 
# row_sums==2 for any restricted to only two habitats...

# filter(m<1|c<1|w<1|r<1)

# This finalizes the table for Table 1

spp_table_data<-btl_sums_patch %>% 
  rename(species=Species) %>% 
  left_join(spp_abund) %>% 
  relocate(n,.after="Guild") %>% 
  mutate(c=case_when(
  c == 0 ~ " ",
  c > 0 ~ "x",
  .default = as.character(c)
)) %>% 
  mutate(m=case_when(
  m == 0 ~ " ",
  m > 0 ~ "x",
  .default = as.character(m)
)) %>% 
  mutate(r=case_when(
  r == 0 ~ " ",
  r > 0 ~ "x",
  .default = as.character(r)
)) %>% 
mutate(w=case_when(
  w == 0 ~ " ",
  w > 0 ~ "x",
  .default = as.character(w)
)) %>% 
  mutate(Guild=case_when(
  Guild == "r" ~ "roll",
  Guild == "t" ~ "tunnell",
  Guild == "d" ~ "dwell",
  .default = as.character(Guild)
)) %>% 
  rename(Species=species,
         N=n,
         Matrix=m,
         Corridor=c,
         Winged=w,
         Rectangular=r)

```

Overall, I collected N = `r total_N` dung beetles (N = `r n_corr` in Connected patches, N = `r n_wing` in Winged patches, N = `r n_rect` in Rectangle patches, N = `r n_mat` in the Matrix). These beetles belonged to N = `r total_spp` species; the N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. *Onthophagus concinnus* was only found in the matrix and winged patches, while *Onthophagus striatulus* was only captured in matrix habitat and rectangular patches. *Geotrupes blackburnii* and *Onthophagus tuberculifrons* were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

### Beetle Abundance

When comparing the overall abundance of beetles (all species combined) across patch types, matrix plots had the highest captures, followed by connected patches, then winged, with the fewest in rectangular patches. Abundances from connected patches were not significantly different from those in the matrix while rectangle and winged patches had significantly fewer than the matrix ($\beta$ = -0.528 and -0.303 respectively, P \< 0.001, Table 4).

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance corresponds to the abundance of *Aphodius alloblackburneus* in matrix patches. Compared to this baseline, results were highly variable, emphasizing species specific responses to patch type. Similarly, species showed to have disproportionate responses to patch since interaction terms varied widely (Table 6).

### Beetle Richness and Diversity

Plotting species richness by patch type reveals consistent richness across patch types with some variation between sampling blocks (Figure 5).The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N). Modeling the effect of patch type on species richness with block as a random effect determined there was no significant differences among patch types. Comparing treatments using matrix patches as a baseline resulted in no significant differences in connected ($\beta$ = 0.216, p = 0.839; Table 7), rectangle ($\beta$ = -0.548, p = 0.583; Table 7), and winged patches ($\beta$ = -0.095, p = 0.663; Table 7).

Biodiversity was also even among patch types, however metric values were varied by sampling blocks. For Shannon Diversity, the estimated mean in matrix patches was 5.202 (SE = 0.665, t = 7.825). None of the alternative patch types showed statistically significant differences compared to matrix: connected ($\beta$ = –0.085, SE = 0.495, t = 0.17), rectangle ($\beta$ = 0.234, SE = 0.495, t = 0.472), or winged ($\beta$ = –0.251, SE = 0.495, t = –0.506). Similarly, for Simpson's Diversity, the average value in matrix patches was 4.181 (SE = 0.640, t = 6.535). Again, none of the other patch types had significant effects: matrix ($\beta$ = -0.107, SE = 0.411, t = -0.259), rectangle ($\beta$ = 0.059, SE = 0.411, t = 0.143), and winged ($\beta$ = –0.480, SE = 0.411, t = –1.169).

\begin{center} 
\textbf{Discussion}
\end{center}
<!-- Paragraph 1: Overview: While there may indeed be effects of patch area and connectivity on dung beetle abundance on community composition, other (landscape-scale) drivers appear to be more important.  -->

This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States. In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches. My main findings emphasized: (1) Habitat type and patch shape were the main driving factors for determining how dung beetle species abundances were composed, however effects were species specific. (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level. (3) Species diversity metrics were also relatively even across patch types however varied widely by sampling blocks. These results suggest that while there may indeed be effects of patch structure and connectivity on dung beetle abundances and community composition, other landscape scale drivers appear to be more prominent for species richness and diversity.

<!-- Paragraph 2: Effects of Edge vs. Connectivity -->

In my comparison of dung beetle compositions, connectivity and habitat edge accounted for differences in dung beetle abundances, yet species richness and diversity were even across patch types and in  the matrix. Total beetle counts were consistently highest in the matrix, in comparison abundances is connected patches were not significantly different while those of the winged and rectangular patches were lower. Additionally, I also observed that patch effects were not equally proportional for all species. This suggests that the species captured could be exhibiting habitat preferences between the open patches versus forested matrix. Another potential explanation is that species in the matrix are acting as a source population which feeds into patches with edge acting as a drift fence for directing movement. The latter example seems more likely because flies, which have similar resource dependencies were also found to have a similar interaction [@fried_habitat_2005], and also studies on dung beetles in the tropics found more dramatic differences in dung beetle populations in fragments and matrix [@WOS:000288810500026]. 
<!-- Paragraph 3: }What then are the drivers of dung beetle diversity and abundance? -->

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition. Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait. Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] 
Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles. In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water. Both of which would effect the land-use by larger mammals [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001]. Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019]. This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.

<!-- Paragraph 4: That said, there were some interspecific differences in abundance.  -->

Species specific differences are apparent but do not follow any particular trend. _Aphodius alloblackburneus_ had a disproportionate positive effect to being captured in matrix patches as compared to other species. Beetles from the genus _Aphodius_ do tend to show patterns of habitat specificity [@roslin_distribution_2001], and many preferring forested habitat*frank*,  so it is not unexpected that they might show a preference towards forested matrix. I did not detect that any species was more positively associated with open patches despite expectations that some species (e.g., _Canthon vigilans_, _Melanocanthon bispinatus_) would prefer open habitat [@conover_phenology_2019; @nealis_habitat_1977]. Altogether this could be another sign pointing towards matrix acting as a source population, and since open habitat was much less dominant in my experimental system beetles could be moving into patches from habitat edge.

<!-- Paragraph 5: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->

Despite ample work documenting their ecological importance, there is a surprising lack of research on dung beetle diversity and corridors outside of the tropics [@nichols_global_2007], with any work in temperate locations coming from Europe where _Aphodius_ beetles dominate [@roslin_distribution_2001]. Many studies in the tropics have found patterns of habitat preference between open fields and continuous forest similar to what I did in this study [@WOS:000349242000004], and some work that found that forest species were using living fences as corridors [@arellano_response_2008]. However, differences in composition were much more distinctly different with some species being exclusive to habitat types unlike I see in this paper. Of the limited work out of the south eastern United States I did recover assemblages similar to other work [@conover_phenology_2019; @nealis_habitat_1977], and even returned results to another study where differences between treatment types were non-significant while variation on a larger scale was more apparent [@young_dung_2023]. In comparison to other studies from SRS my results were similar to those of small mammals which were more likely to be using matrix land [@mabry_influence_2003]. I did expect some species to be more dominant in patches with higher edge ratios as was found with ants at SRS [@resasco_landscape_2014], but species were evenly present in all patches.

<!-- Paragraph 6: Caveats -->

There were some limitations of this study which should be addressed in any future work to capture a clearer picture of how populations are being altered by habitat connectivity. First, sampling for this study was conducted over a two month period in the summer of 2024, sufficient data was collected but due to a lack of available resources temporal patterns were obscured due to inconsistent sampling periods. This is highly important since dung beetles exhibit consistent patterns of seasonality [@conover_phenology_2019; @davis_feeding_1966]. Another potential avenue for improvement is lowering the grain size of sampling and changing trap placement to better understand the effects of edge proximity and connectivity. I also used one of the best bait types for collecting dung beetles in pig dung [@marsh_optimising_2013], but a mix of differently sourced baits may have been more optimal as more diverse baits attract more diverse species [@frank_nutrient_2017; @frank_search_2018; @gimenez_gomez_dung_2021].

<!-- Paragraph 7: Moving forward -->

Moving forward, there is a bounty of knowledge yet to be collected for dung beetles in the south eastern United States. Relative to what is already known in tropical ecosystems, this study provides a glimpse on how dung beetles are responding to fragmentation and connectivity in subtropic pine habitats, but much can still be learned about dung beetles in these spaces and the role they play in ecosystems as well as how they interact with landscapes. The unique role of dung beetles as waste removers is fascinating enough and of great importance for maintaining landscapes so dominated by pastures like Southern US [@cheng_dweller_2022; @WOS:000368243300001]. Not only waste removal but also secondary effect such as seed dispersal and parasitic reduction are important to understand in the context of fragmentation [@fincher_effects_1975; @vulinec_dung_2002]. SRS provides an excellent experimental design for direct comparison of landscapes but another main goal of the site is to study dispersal, dung beetles would be an excellent system for studying movement. They already show promise, in preliminary trials I released beetles from both the connected and rectangular patches with baited traps in the central patch. I recovered one recapture which originated from the connected patch, indicating that individuals do move through corridors.

<!-- Paragraph 8: Close with takeaways for conservation and management -->

Although there is plentiful work to be done, this study paints an encouraging picture that while dung beetles face disturbances that make habitats less preferential, they are robust enough to persist throughout fragmented landscapes and the management strategies we do have to connect fragmented landscapes work to mitigate losses. Because assemblages remain mostly the same across these landscapes, ecosystem services should remain uninterrupted allowing continuing benefits to all community members. While this study of dung beetles in fragmented landscapes was conducted in a relatively protected area, in real world application we could consider the use of movement corridors within disturbed areas to help bolster effected populations, but also near areas of high importance such as pastures to provide areas of refuge for beetles. In this case it is important to consider the inverse of what was manipulated in this study where open field acts as matrix and forested area is focal landscape, thus it is vital to continue learning more about how landscape composition and connectivity effect dung beetle populations.

\newpage

\begin{center} 
\textbf{References}
\end{center}

\smallskip

::: {#refs}
:::

\newpage

<!-- (text - Nathan Fig of plots) -->

\newpage

<!-- ```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE,fig.cap = 'Experimental landscapes indicating location of the pitfall traps.',  out.width="25%", fig.align='center'} -->

<!-- a<-readPNG("images/pitfalls.png") -->

<!-- b<-readPNG("images/srs-map.png") -->

<!-- grid.arrange(rasterGrob(a),rasterGrob(b),ncol=1) -->

<!-- ``` -->

```{r echo=FALSE, fig.align='center', fig.cap="All experimental blocks at SRS. Filled squares indicate the blocks in which dung beetles were sampled. Each block has a unique ID number.", message=FALSE, warning=FALSE, label=srs-map, out.width="60%"}

knitr::include_graphics("images/srs-map.png")

```

```{r, label = patches, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Overhead view of one of the experimental blocks indicating the different types of patches and the distance between them (from Nathan et al. 2018 and reused with permission).", out.width="70%", fig.align='center'}

knitr::include_graphics("images/patches-pic.png")

```

```{r, label = pitfalls, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Diagram indicating the placement pitfall traps in each patch and in the matrix. Traps in patches were approximately 40 m from the patch edge.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/pitfalls.png")

```

<!-- ```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- library(iNEXT) -->

<!-- # -->

<!-- # btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->

<!-- # spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->

<!-- btl_sums_patch<-btl_data %>% -->

<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->

<!--   group_by(species,patch,block) %>% -->

<!--   mutate(patch=case_when( -->

<!--     patch == "c" ~ "Corridor", -->

<!--     patch == "m" ~ "Matrix", -->

<!--     patch == "w" ~ "Winged", -->

<!--     patch == "r" ~ "Rectangle", -->

<!--     .default = as.character(patch))) %>% -->

<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->

<!--   arrange(patch,desc(n)) %>% -->

<!--   rename(sp_code=species, -->

<!--          n_patch=n) %>% -->

<!--   mutate(pb=paste(patch,block,sep="-")) %>%  -->

<!--   ungroup() %>%  -->

<!--   select(-patch,-block) %>%  -->

<!--   pivot_wider( -->

<!--     names_from = pb, -->

<!--     values_from = c(n_patch))  -->

<!-- div.data <- btl_sums_patch |> -->

<!--   column_to_rownames(var = "sp_code") -->

<!-- div.btl <- iNEXT(div.data, q = c(0,1,2), datatype = "abundance") -->

<!-- z <- iNEXT(div.data, q=c(0,1,2), datatype="abundance") -->

<!-- ggiNEXT(z, facet.var="Order.q", color.var="Assemblage") + -->

<!--   theme_bw(base_size=18) -->

<!-- ``` -->

\newpage
```{r, label = patch, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average total dung beetle abundance by patch type with standard deviations as error bars.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/total_abund.png")

```

```{r, label = abundance, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average abundance of the top 6 most abundant species by patch type.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/sp_abund_patch.png")

```

```{r, label = richness, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle species richness in each patch type. The point shapes indicates the block in which each patch was located.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/sp_richness_patches.png")
```

```{r, label = shannon, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Shannon diversity by patch type.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/sp_shannon_patches.png")

```

```{r, label = simpson, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Dung beetle Simpson's index y patch type.", out.width="70%", fig.align='center'}

knitr::include_graphics("images/sp_simpson_patches.png")

```

<!-- ```{r sp_richness_patches, echo = FALSE, fig.cap = "Dung beetle secies richness in three different --- and the forest matrix surrounding patches.", out.width = '100%'} -->

<!-- knitr::include_graphics("./corridor_beetles/corridor_docs/eric_ms_thesis/images/sp_richness_patches.png") -->

<!-- ``` -->

\newpage

```{r summary-table, echo=FALSE,message = FALSE,warning=FALSE}
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

spp_table<-knitr::kable(spp_table_data, 
      digits = 2,
      align="lcrcccc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Dung beetle species sampled in the SRS site and their total abundance over the course of the study.") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>% 
  column_spec(1, italic = T) %>% 
  row_spec(0, bold = T) 
# %>% 
  # row_spec(1:15, hline_after = T) 
spp_table

```

\newpage

<!-- ```{r abundance-patch, echo=FALSE,message = FALSE,warning=FALSE} -->
<!-- # https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html -->

<!-- btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv")) -->
<!-- spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv")) -->

<!-- btl_sums_patch<-btl_data %>% -->
<!--   select(-sum) %>% -->
<!--   pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% -->
<!--   group_by(patch) %>% -->
<!--   summarize(n=sum(n, na.rm=TRUE)) %>% -->
<!--   arrange(patch,desc(n)) %>% -->
<!--   mutate(patch=case_when( -->
<!--     patch == "c" ~ "Corridor", -->
<!--     patch == "m" ~ "Matrix", -->
<!--     patch == "w" ~ "Winged", -->
<!--     patch == "r" ~ "Rectangle", -->
<!--     .default = as.character(patch))) -->

<!-- # if wanted to make a bar graph -->
<!-- # ggplot(btl_sums_patch, aes(x=patch, y=n)) +  -->
<!-- #   geom_bar(stat = "identity") -->
<!-- #  -->
<!-- abund_patch <- knitr::kable(btl_sums_patch, -->
<!--       digits = 2, -->
<!--       align="lr", -->
<!--       format="latex", -->
<!--       row.names = FALSE, -->
<!--       booktabs=T, -->
<!--       linesep = "", #removes the blank line after every 5 lines -->
<!--       caption = "Total dung beetles captured in all replicates of a patch type.") %>% -->
<!--   kable_styling(bootstrap_options = c("striped"), -->
<!--                 full_width = F, -->
<!--                 latex_options=c("scale_down","hold_position"), -->
<!--                 font_size = 12, -->
<!--                 position = "center") %>% -->
<!--   # column_spec(1, italic = T) %>% -->
<!--   row_spec(0, bold = T) -->
<!--   # row_spec(1:15, hline_after = T) -->
<!-- abund_patch -->

<!-- ``` -->

\newpage

```{r bray-curtis, echo=FALSE,message = FALSE,warning=FALSE}
btl_spec_patch<- btl_data |>
  pivot_longer(pvin:osyl, names_to = "species", values_to = "n") |>
  group_by(species,patch) |>
  summarise(n=sum(n, na.rm=TRUE)) |>
  arrange(patch,desc(n)) |>
  pivot_wider(values_from=n,names_from=species) |>
  mutate(patch=case_when(
    patch == "c" ~ "Corridor",
    patch == "m" ~ "Matrix",
    patch == "w" ~ "Winged",
    patch == "r" ~ "Rectangle",
    .default = as.character(patch))) |>
  column_to_rownames("patch")

taxa_bray <- vegan::vegdist(btl_spec_patch, method = "bray", binary = FALSE)

# taxa_bray
tb<-as.matrix(taxa_bray)
tb[lower.tri(tb)] <- 0
tb<-as.data.frame(tb)
tb<-tb %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  select(-Corridor)

row.names.remove <- c("Winged")
tb<-tb[!(row.names(tb) %in% row.names.remove), ]
tb[tb == 0] <- ""
col1<-row.names(tb)
col1<-as.data.frame(col1)

tb<-bind_cols(col1,tb) %>% 
  rename(" "=col1)
# bray<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/bray_patches.csv") 
# bray<-read_csv(here("corridor_docs","eric_ms_thesis","tables","bray_patches.csv")) %>% 
  bray<-tb %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Bray-Curtis dissimilarity comparing composition of dung beetles between patches.") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T) %>% 
  column_spec(1, bold = T)
  # row_spec(1:15, hline_after = T)
bray
```

\newpage

<!-- ```{r Table-aovMO, echo=FALSE,message = FALSE,warning=FALSE} -->
<!-- aovM0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM0.csv")) %>% -->
<!--   mutate(term=case_when( -->
<!--     term == "patch_type" ~ "Patch type", -->
<!--     .default = as.character(term))) |> -->
<!-- # aovM0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM0.csv") %>%  -->
<!--   kable( -->
<!--       digits = 3, -->
<!--       align="lcc", -->
<!--       format="latex", -->
<!--       row.names = FALSE, -->
<!--       booktabs=T, -->
<!--       linesep = "", #removes the blank line after every 5 lines -->
<!--       caption = "AOV TABLE CAPTION") %>% -->
<!--   kable_styling(bootstrap_options = c("striped"), -->
<!--                 full_width = F, -->
<!--                 latex_options=c("scale_down","hold_position"), -->
<!--                 font_size = 12, -->
<!--                 position = "center") %>% -->
<!--   # column_spec(1, italic = T) %>% -->
<!--   row_spec(0, bold = T) -->
<!--   # row_spec(1:15, hline_after = T) -->
<!-- aovM0 -->
<!-- ``` -->

\newpage
\blandscape
```{r Table-MO, echo=FALSE,message = FALSE,warning=FALSE}
M0<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M0.csv")) %>%
  replace_na(list(group=" ")) %>% 
  rename(SE = std.error,
         Estimate=estimate,
         Term=term,
         Group=group,
         Effect=effect,
         Statistic=statistic,
         P=p.value) %>% 
  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
  mutate(Group=if_else(Group=="block","Block",Group)) %>% 
  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
  mutate(Term=gsub("[.]"," x ",Term)) %>% 
  mutate(SE=round(SE,3)) %>% 
  mutate(Estimate=round(Estimate,3)) %>% 
  mutate(Statistic=round(Statistic,3)) %>% 
  mutate(P=round(P,3)) %>% 
  mutate_all(as.character) %>% 
  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
  replace_na(list(SE="",Statistic="",P="")) %>%

  # need to round numbers above to 4, convert to character, then na_replace
  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
# M0<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M0.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Model of species abundance by patch type.") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M0
```
\elandscape
\newpage

```{r Table-aovMO3, echo=FALSE,message = FALSE,warning=FALSE}
aovM3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","aovM3.csv")) %>%
    mutate(term=case_when(
    term == "sp_code" ~ "Species",
    term == "patch_type" ~ "Patch type",
    term == "sp_code:patch_type" ~ "Species x Patch type",
    .default = as.character(term))) |>
  mutate(p.value=as.character(p.value)) %>% 
  mutate(p.value=if_else(as.numeric(p.value)<0.001,"< 0.001",p.value)) %>% 
  rename(P=p.value,
         Term=term,
         Statistic=statistic) %>% 
# aovM3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/aovM3.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "AOV patch type effects on species abundance and interavtive term.") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
aovM3
```

\newpage

\blandscape
```{r Table-M3, echo=FALSE,message = FALSE,warning=FALSE}
M3<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M3.csv")) %>%
  replace_na(list(group=" ")) %>% 
  # mutate(p.value=as.character(p.value)) %>% 
  # mutate(p.value=if_else(as.numeric(p.value)<0.001,"< 0.001",p.value)) %>% 
  rename(P=p.value,
         SE = std.error,
         Estimate=estimate,
         Term=term,
         Group=group,
         Effect=effect,
         Statistic=statistic) %>%

  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
  mutate(Group=if_else(Group=="block","Block",Group)) %>% 
  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
  mutate(Term=gsub("sp_code","Species - ",Term)) %>% 
  mutate(Term=gsub("[.]"," x ",Term)) %>% 
  mutate(Term=gsub("[:]"," : ",Term)) %>% 
  mutate(SE=round(SE,3)) %>% 
  mutate(Estimate=round(Estimate,3)) %>% 
  mutate(Statistic=round(Statistic,3)) %>% 
  mutate(P=round(P,3)) %>% 
  mutate_all(as.character) %>% 
  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
  replace_na(list(SE="",Statistic="",P="")) %>%

  # need to round numbers above to 4, convert to character, then na_replace
  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
# M3<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M3.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "M3 patch type effects on species abundance of top 6 most abundant speceies and interactive term for patch type and species ID.") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M3
```
\elandscape
\newpage

```{r Table-M-rich, echo=FALSE,message = FALSE,warning=FALSE}
M_rich<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_rich.csv")) %>%
   mutate(term=case_when(
    term == "patch_typeMatrix" ~ "Patch Type - Matrix",
    term == "patch_typeRectangle" ~ "Patch Type - Rectangle",
    term == "patch_typeWinged" ~ "Patch Type - Winged",
    term == "patch_typeConnected" ~ "Patch Type - Connected",
    .default = as.character(term))) |>
  rename(P=p.value,
         SE = std.error,
         Estimate=estimate,
         Term=term,
         Effect=effect,
         Statistic=statistic) %>%

  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>%

  mutate(SE=round(SE,3)) %>% 
  mutate(Estimate=round(Estimate,3)) %>% 
  mutate(Statistic=round(Statistic,3)) %>% 
  mutate(P=round(P,3)) %>% 
  mutate_all(as.character) %>% 
  mutate(P=if_else(as.numeric(P)<0.001,"< 0.001",P)) %>% 
  replace_na(list(SE="",Statistic="",P="")) %>%

# M_rich<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_rich.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Richness Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_rich
```

\newpage

```{r Table-M-shan, echo=FALSE,message = FALSE,warning=FALSE}
M_shan<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_shan.csv")) %>%
  replace_na(list(group=" ")) %>%

  rename(SE = std.error,
         Estimate=estimate,
         Term=term,
         Group=group,
         Effect=effect,
         Statistic=statistic) %>%

  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
  mutate(Group=if_else(Group=="block","Block",Group)) %>%

  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
  mutate(SE=round(SE,3)) %>% 
  mutate(Estimate=round(Estimate,3)) %>% 
  mutate(Statistic=round(Statistic,3)) %>%

  mutate_all(as.character) %>%

  replace_na(list(SE="",Statistic="")) %>%

  # need to round numbers above to 4, convert to character, then na_replace
  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
# M_shan<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_shan.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Shannon Diversity Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_shan
```

\newpage

```{r Table-M-simp, echo=FALSE,message = FALSE,warning=FALSE}
M_simp<-read_csv(here("corridor_docs","eric_ms_thesis","tables","M_simp.csv")) %>%
  replace_na(list(group=" ")) %>% 
  rename(SE = std.error,
         Estimate=estimate,
         Term=term,
         Group=group,
         Effect=effect,
         Statistic=statistic) %>%

  mutate(Effect=if_else(Effect=="ran_pars","Random",Effect)) %>% 
  mutate(Effect=if_else(Effect=="fixed","Fixed",Effect)) %>% 
  mutate(Group=if_else(Group=="block","Block",Group)) %>%

  mutate(Term=gsub("patch_type","patch type - ",Term)) %>% 
  mutate(Term=gsub("cor__","correlation: ",Term)) %>% 
  mutate(Term=gsub("sd__","SD: ",Term)) %>% 
  mutate(SE=round(SE,3)) %>% 
  mutate(Estimate=round(Estimate,3)) %>% 
  mutate(Statistic=round(Statistic,3)) %>%

  mutate_all(as.character) %>%

  replace_na(list(SE="",Statistic="")) %>%

  # need to round numbers above to 4, convert to character, then na_replace
  # replace_na(list(group=" ", std.error="",statistic=" ",p.value = "")) %>% 
# M_simp<-read_csv("~/Rstudio/corridor_beetles/corridor_docs/eric_ms_thesis/tables/M_simp.csv") %>% 
  kable(
      digits = 3,
      align="lcc",
      format="latex",
      row.names = FALSE,
      booktabs=T,
      linesep = "", #removes the blank line after every 5 lines
      caption = "Simpson's Index Model") %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                latex_options=c("scale_down","hold_position"),
                font_size = 12,
                position = "center") %>%
  # column_spec(1, italic = T) %>%
  row_spec(0, bold = T)
  # row_spec(1:15, hline_after = T)
M_simp
```

\newpage

\vspace*{1cm}

\begin{center}
\textbf{BIOGRAPHICAL SKETCH}
\end{center}

<!-- _A biographical sketch is required of all candidates. The biographical sketch should be in narrative form. Third person, past tense, it typically includes the educational background of the candidate. The author should have replaced this paragraph with their own._ -->

\doublespacing

Eric Escobar-Chena completed his Bachelors education at Virginia Commonwealth University in 2023. During his time there he developed a fondness for insects which grew into a curiosity of the natural world. He later began to explore this curiosity deeper in beginning his graduate education at the University of Florida as a Master's Student under the supervision of Emilio Bruna.