### 2025-08-02 corridor_docs/2024_corridor_mtg_notes.Rmd

---
title: "Corridor Meeting"
author: "Emilio M. Bruna"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTORY OVERVIEW

## Nick Haddad
### Overview of the Corridor Expeoriment

**Capacity Building**

-   79 undergrads
-   39 postbac scientists
-   27 grad students
-   9 postdocs

**Overarching Question**

Do corridors work for dispersal (phase 1) - Builds on a priort project
investigating the effects of corridor *length* - Corridor project
originally designed around 1 sp of butterfly (spicebush swallowtail) -
corridors and wind - maybe use this to ask questions about volatiles and
detection? - spicebush swallowtail only species that *didn't* use
corridors

Do corridors increase diversity (phase 2) - 25 years of plant diversity
surveys - effect of fire ants on other ant species - spiders

Plant population dynamics and species interactions - restoration plants
and interactions (with Lauren Sullivan MSU) - observed depredation of
butterflies, maybe by prayinf mantis?

Outreach - Conservation Corridor - research digests - 30 x 30
initiatives

EB notes: reminds me very much of fragmentation work

-   interactions & community composition, jump over population dynamics
-   limited matrix comparisons

## Lars Brudvig

### The Remnant Project

**Overarching question** How best to restore longleaf pine remnants?

-   At SRS in 1951 most of LLP forest was cleared for ag, but there were
    remnants, perhaps in woodlots
-   this was followed by the largest planting effort in the world, with
    planted forest and interspersed patches of llp
-   remnant project: a remnant of forest adjacent to an ag field that
    was abandoned and replanted (post ag).
-   each of remnant and post-ag site had two plots, with in the post-ag
    one being thinned and the other being unthinned
-   Questions:
    -   What are consequences of past ag land use?
    -   what are consequences of restoration thinning treatment
    -   can thinning ameliiorate legacies of ag land use?
    -   to what extent can remnants serve as sources for restoration and
        post-ag land areas?

-   Lesson 1: effects of ag history are long-lived
    -   ag history alters plant
        composition.
    -   about 60 spp of plants only foundin remnant fragments even
        though there have been decades for them to move only meters into
        adjacent restoration areas
    -   ag legacy on soils
-   Lesson 2
    - (missed)

-   Lesson 3: what limits plant colonization post ag
    - limited seed production? In absence of thinning low flowering and seed production, much higher seed production
    - limited seed dispersal? most of occurrence in post-ag is in seed additiona plots, with more establishment in thinning. def some limitation.
    - establishment / persistence
    - many plant species will manually resestablish when manually added

## Ellen Damschen
### SERDP Project: restoring the ground layer community

- Brett Mattingly, John Orrick, Ellen, and Lars
- thinking about gradients in fire history 
- study also at ft stuart and fort liberty (formerly ft bragg)
- how to most effectively do restoration?
- asking about role of adding seeds relative to a bunch of other processes? 
    - consumer exclusion
    - elimination of competition by spraying with herbicide
- initial: 108 sites, 30 million seeds of 25 species, veg data 2010-2013

- seed addition consistently increases diversity over time
- other ket factords seed coat, liiter depth (fire burns off), temp 
- once established, recovered populations persist and spread
- management can affect recovery and spread
    - prescribed fire and thinning
    - adding seeds

# STUDENT PRESENTATIONS

## Melissa Burt (did PhD with Susan Whitehead at VTech)
### Ants

- >144K ants identified, 60+ species
- 45% fire ants
- spp richness increasing over time everywhere, at higher rate in connected (effect of fire ants)
- patches with corridors have fewwr species than isolatoned and high edge patches
- fewer species 
EB note - has dung beetle bycatch in pitfalls

## Katherine
### PLant arthropod interactions

- how do different factors influence multiple plant-arhropod interactions that can influence reproductive output: pollination, florivory, and depredation of pollinators by spiders
- how do these factors influence pollination rate?

used a founder species (_Carphephorus bellidifolius_)
founder species = species inplanted in 2007(?) experiment where introduced into different patches in edge-interior grid pattern

- observes pollinator visits for 10 min
- collect folivores and spiders _after_ the 10 min visit
- surveyed floral abundnace within 5 m of plant
- collected seeds ar end of season to measure pollination rate

## Thomas
### Edge Effects, matrix, and spillover

- edge effects and spilloiver
- population density leads to spillover
- cons of spillover - disease, predators, etc.
- 2009 (PNAS) study by lars et al - spillover out of connected patches into matrix greater, edge dependent decreases
- spillover driver by diversity at the patch level
- shortly after - spillover effect
- 2011 matrix thinned for red cockaded woodpecker

## Angelica Bautista
### Seed emergence and establishment in LLP